A Container-Based Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-10-16 draft.tex
Update draft.tex

After starting the application, users can push several Experiment
Definition files and start monitoring the current execution in the
terminal.  After an experiment is finished, the user must execute
another script that takes as argument the experiment id to generate a
collection of files containing all the data generated by the
experiment, again in a JSON format. An example of an experiment
definition is shown in Listing \ref{code:exp}. We give more details
about the data contained in the file in the next section. Users can
use these files to analyze and plot experiment data. We include Python
scripts in the repository to plot the running times for the
experiments and all other charts used in this paper; there is also a
script to generate the files needed by the COCO framework
\cite{hansen2016coco}, which can generate standard comparisons against
other methods.

@@

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. This section gives specific details
about how we deployed EvoSwarm in a cloud setting for running the experiments
within this paper. Running the cloud-native application in the cloud has the obvious
benefit of a multi-instance deployment, where we can run containers in several
computing nodes at the same time, not only in a single machine as it is the case when using docker-compose.

@@

We chose the second option, a container orchestration service. We
deployed the EvoSwarm application in Amazon Elastic Container Service (ECS)
since it needs fewer changes in configuration (from deploying it locally using
{\tt docker-compose}) and it has a lower cost than the Kubernetes
alternative. This does not imply any loss of generality; first,
because it's {\tt docker-compose} compatible, and it can be deployed
in approximately the same way to any of the services, mentioned above,
compatible with it. Since {\tt docker-compose} is open source, it can
be ported to any computing node, and can thus be deployed to any other
cloud service like Azure or Google Cloud Platform. At any rate, please
note that there are several levels of infrastructure programming
here. At the lowest level, we are using docker containers, which are
compatible across a myriad of cloud providers in one way or other. At
the next level, we are using a container composition technology that
is also standard, {\tt docker compose}, once again compatible across
any platform that can host that kind of service. There's, however, a
final, platform-specific, infrastructure definition that will have to
vary across any public or private cloud platform that we are using,
and that will tap the cloud provider API. These changes, however, are
usually minimal, and can also be done through the CLI or using the web
console.

@@

The complete task configuration is generated by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example, to deploy the controller, setup, and Redis 
containers we used the file in Listing ~\ref{code:compose}.

Each container is defined as a {\texttt service} ({\em task} in AWS terms), with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\footnote{They are shown as different repositories for user {\tt
    mariosky} here: \url{https://hub.docker.com/u/mariosky}}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally, {\texttt environment}
variables used by code running inside containers.

@@

Figure \ref{fig:aws-configuration} shows the basic EC2 configuration
used for the deployment of experiments. We selected three types of EC2
instances according to our needs; the {\texttt setup}, {\texttt
controller}, and {\texttt redis} containers do not demand many
resources and we do not need to scale them as the amount of parallel
workers are increased. They run in a single {\texttt C5 High-CPU
large} instance. With the available 2048 CPU shares divided as
follows: {\texttt setup} (128), {\texttt controller} (896), {\texttt
redis} (1024). With this assignment, the {\texttt redis} container
gets a dedicated vCPU, needed because it is the most demanding task of
the three. We assigned 2048 units to each {\texttt worker} container,
and using a more capable {\texttt C5 High-CPU XLarge} instance, with
four vCPUs and 4096 CPU shares, with this configuration each worker
gets a pair of vCPUs.  We scaled the number of instances as needed,
using one instance for every two {\texttt worker} containers.

Because all these instances run in a private subnet, we also run a
Bastion instance, that is, a specially protected network node, in a
public subnet, from where we send the commands to execute the
experiments and later collect the results. We connected to this
instance through a {\texttt ssh} session from our personal
workstation. This is a {\texttt R5AD Large} instance, selected with
75GiB of storage, needed for keeping the logs, and more memory (16GiB)
for processing the logs in memory. The details these EC2 instances and
their cost at the time of writing are shown in Table \ref{tab:ec2}.

@@

The complete configuration files are available at public GitHub repository 
\footnote{\url{https://github.com/mariosky/EvoSwarm/tree/master/aws-EC2}}, and
they are used in the next section to carry an extensive evaluation of the framework.
Again, each configuration describes the deployment of containers in a consistent
and portable way, and it was used in each phase of the life cycle, from
development through testing to production independently on the kind of
infrastructure.

@@

In the previous section, we have described an option for an EvoSwarm
deployment that can be used by the public to create reproductive
research in the cloud; in this section, we are using the same
framework to learn if the proposed solution can efficiently improve
the scalability and performance of population-based optimization
algorithms.  Hence, in the following sections, we provide answers to
the following questions:

\begin{itemize}
\item Can we improve the execution time of the algorithm by adding populations
 and serverless functions (workers)  to the system and, in particular,
 what is its effect on the scalability of the system? % Improve with respect to what? Single population? 
                         %Please explain carefully the issues and how scaling is related to the number of populations. - JJ
                         % The question could be that, How scaling is related to number of populations and workers in the system ? - M

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increase the performance of the
search by needing fewer function evaluations than a homogeneous setting? 
% Moved reproducibility question from here - M

>>>>

After starting the application, users can push several Experiment
Definition files and monitor the current execution using the PC console.
After an experiment is finished, the user must execute
another script that takes as argument the experiment id to generate a
collection of files containing all the data generated by the
experiment, again in a JSON format. An example of an experiment
definition is shown in Listing \ref{code:exp}. We give more details
about the data contained in the file in the next section. Users can
use these files to analyze and plot experiment data. We include Python
scripts in the repository to plot the running times for the
experiments and all other charts used in this paper; there is also a
script to generate the files needed by the COCO framework
\cite{hansen2016coco}, which can generate standard comparisons against
other methods.

@@

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. This section gives specific details
about how we deployed EvoSwarm in a cloud setting for running the experiments
within this paper. Running the cloud-native application in the cloud has the obvious
benefit of a multi-instance deployment. We can run containers in several
computing nodes at the same time, not only in a single machine as it is the case when using docker-compose.

@@

We chose the second option, a container orchestration service. We
deployed the EvoSwarm application in Amazon Elastic Container Service (ECS)
since it needs fewer configuration changes (from deploying it locally using
{\tt docker-compose}) and it has a lower cost than the Kubernetes
alternative. This does not imply any loss of generality; first,
because it's {\tt docker-compose} compatible, and it can be deployed
in approximately the same way to any of the services, mentioned above,
compatible with it. Since {\tt docker-compose} is open source, it can
be ported to any computing node, and can thus be deployed to any other
cloud service like Azure or Google Cloud Platform. At any rate, please
note that there are several levels of infrastructure programming
here. At the lowest level, we are using docker containers, which are
compatible across a myriad of cloud providers in one way or other. At
the next level, we are using a container composition technology that
is also standard, {\tt docker compose}, once again compatible across
any platform that can host that kind of service. There's, however, a
final, platform-specific, infrastructure definition that will have to
vary across any public or private cloud platform that we are using,
and that will tap the cloud provider API. These changes, however, are
usually minimal, and can also be done through the CLI or using the web
console.

@@

The complete task configuration is generated by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example, to deploy the controller, setup, and Redis 
containers, we used the file in Listing ~\ref{code:compose}.

Each container is defined as a {\texttt service} (a {\em task} in AWS terms), with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\footnote{They are shown as different repositories for user {\tt
    mariosky} here: \url{https://hub.docker.com/u/mariosky}}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally, {\texttt environment}
variables used by code running inside containers.

@@

Figure \ref{fig:aws-configuration} shows the basic EC2 configuration
used for the deployment of experiments. We selected three types of EC2
instances according to our needs; the {\texttt setup}, {\texttt
controller}, and {\texttt redis} containers do not demand many
resources and we do not need to scale them as the number of parallel
workers is increased. They run in a single {\texttt C5 High-CPU
large} instance. With the available 2048 CPU shares divided as
follows: {\texttt setup} (128), {\texttt controller} (896), {\texttt
redis} (1024). With this assignment, the {\texttt redis} container
gets a dedicated vCPU because it is the most demanding task.
We assigned 2048 units to each {\texttt worker} container
and used a more capable {\texttt C5 High-CPU XLarge} instance, with
four vCPUs and 4096 CPU shares. With this configuration, each worker
gets a pair of vCPUs.  We scaled the number of instances as needed,
using one instance for every two {\texttt worker} containers.

Because all these instances run in a private subnet, we also run a
Bastion instance, a specially protected network node, in a
public subnet, from where we send the commands to execute the
experiments and later collect the results. We connected to this
instance through an {\texttt ssh} session from our workstation.
This is an {\texttt R5AD Large} instance, selected with
75GiB of storage, needed for keeping the logs, and more memory (16GiB)
for processing the logs in memory. The details of these EC2 instances and
their cost at the time of writing are shown in Table \ref{tab:ec2}.

@@

The complete configuration files are available at public the GitHub repository 
\footnote{\url{https://github.com/mariosky/EvoSwarm/tree/master/aws-EC2}}, and
they are used in the next section to carry an extensive evaluation of the framework.
Again, each configuration describes the deployment of containers in a consistent
and portable way. It was used in each phase of the life cycle, from
development through testing to production independently on the kind of
infrastructure.

@@

In the previous section, we have described an option for an EvoSwarm
deployment that can be used by the public to create reproductive
research in the cloud; in this section, we are using the same
framework to learn if the proposed solution can efficiently improve
population-based optimization algorithms' scalability and performance.
Hence, in the following sections, we provide answers to
the following questions:

\begin{itemize}
\item Can we improve the algorithm's execution time by adding populations
 and serverless functions (workers)  to the system and, in particular,
 what is its effect on the scalability of the system? % Improve with respect to what? Single population? 
                         %Please explain carefully the issues and how scaling is related to the number of populations. - JJ
                         % The question could be that, How scaling is related to number of populations and workers in the system ? - M

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increase the search's
performance by needing fewer function evaluations than a homogeneous setting? 
% Moved reproducibility question from here - M
